---
title: "Groundwater model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groundwater-model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load-package}
library(nitratesgame)
```


<!-- --- -->
<!-- title: "Groundwater model" -->
<!-- author: "Gopal Penny" -->
<!-- date: "10/27/2020" -->
<!-- output:  -->
<!--   html_notebook: -->
<!--     toc: true -->
<!--     number_sections: true -->
<!--     code_folding: hide -->
<!-- --- -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(sf) # needed for package to represent well capture polygons
library(units) # needed for package, ensures proper units
library(ggforce) # for plotting axes with units
library(patchwork) # for aligning axes of multiple subplots
```

# Groundwater model

First we set some basic parameters for the scenario including:

* Household water use
* Seepage rate
    + Annual precipitation
    + Seepage fraction
* Seepage fraction
* Well source area
    + radius of source area
* Housing density

```{r run_gw_model, message = FALSE}
# Get area of hh withdrawal
# hh_annual <- set_units(76, "gallon/day") %>% # from USGS water data
#   set_units("ft^3/year") * 4
# precip_annual <- set_units(1070, "mm / year") #%>%
# seepage_fraction <- 0.4
# seepage_annual <- precip_annual * seepage_fraction
# area_of_withdrawal <- hh_annual / seepage_annual

# wells are screened from 10 ft to 20 ft depth below the water table
z1 <- set_units(10,"ft")
z2 <- set_units(20,"ft")

rs <- set_units(20, "ft") # this is the radius of the region of well capture
density <- set_units(0.5, "1/acre") # the housing density
```

Now we need a grid of septic systems. You can create an array of wells yourself or use `get_hh_grid`. The inputs to `get_hh_grid` are `density` and `area` as units objects.


```{r}
area <- set_units(64, "acre")
hh_array <- get_hh_grid(density, area)
hh_array$id <- 1:nrow(hh_array) # supply an id
```

The groundwater model predicts the probability that a point source will contaminate at least one well in an array of wells. However, in this case we are interested in the probability that an array of septic fields (considered point sources) will lead to contamination in a single well. This problem can be modeled using `get_intersection_probability` by considering the well as a point source and the septic systems as "virtual wells" such that the spatial relationship between a well and septic system (point source) is the same as the relationship between the virtual well and well (now considered the point source). The function `get_septic_well_array` does this job for us. See the `?get_septic_well_array` for details.

```{r}
virtual_well_array <- get_septic_well_array(hh_array, "septic", z_range = c(z1, z2), rs = rs)
```

Here you can see that the virtual well array is identical to the septic array but rotated by 180 degrees around the origin.

```{r fig.width = 6}
library(ggplot2)
library(ggforce) # needed to plot axes using units objects
ggplot(mapping = aes(x, y, color = id)) +
  geom_point(data = hh_array, aes(shape = "septic systems"), size = 4, stroke = 1) +
  geom_point(data = virtual_well_array, aes(shape = "virtual wells"), size = 2) +
  scale_shape_manual(values = c(1, 16)) +
  scale_color_viridis_c(option = "B") + coord_equal()
```

The array `virtual_well_array` contains sf geometries for each of the wells in $\theta-\phi$ space. We can plot these geometries projected onto the z-axis such that it appears as viewed from the point source at the land surface. To do so we can take $z_{projected}=\sin(\phi)$ and plot using `geom_rect` from `ggplot`. We zoom in only on $\theta \in [0,pi/2]$.

```{r fig.width = 6}
p_wells <- ggplot(virtual_well_array) + 
  geom_rect(aes(xmin = theta1, xmax = theta2, ymin = sin(phi1), ymax = sin(phi2), fill = id), color = NA, alpha = 1) +
  scale_fill_viridis_c("id") +
  scale_y_reverse("sin(phi)") + 
  scale_x_continuous(breaks = c(0, pi/2), labels = c("0", "pi/2")) +
  coord_cartesian(xlim = c(0, pi/2))

p_phi <- ggplot(data.frame(alpha = seq(0,20,by=.1)) %>% dplyr::mutate(phi = atan(1/alpha))) + 
  stat_density(aes(sin(phi))) +
  scale_x_reverse("sin(phi)") + 
  coord_flip()
p_wells + p_phi + patchwork::plot_layout(widths=c(0.8,0.2))
```

## Get probability

Now get the probability of contamination of the well, using the virtual well array.

```{r}
gw_example <- virtual_well_array %>%
  get_intersection_probability(theta_range = c(0,pi/2), alpha_range = c(0, 20),
                               self_treat = FALSE, show_progress = FALSE)
gw_example
```

## Get multiple probabilities at once

Now, let's get probabilities for each row of a `tibble` which contains columns for all of the appropriate input variables. Note that this must be a `tibble` and not a `data.frame`, because it has to store complex objects (i.e., `list` columns).

```{r}
# Create a single row with constant parameters
params_row <- tibble(
  z1 = set_units(10, "ft"),
  z2 = set_units(20, "ft"),
  area = set_units(64, "acre"),
  theta_range = list(c(0, pi/4)), # this will be unlisted in the function
  alpha_range = list(c(0, 20)), # this will be unlisted in the function
  hh_array_type = "septic")

# Add varying parameters
params_df <- params_row %>%
  crossing(density = set_units(c(0, 0.5), "1/acre"), 
           rs = set_units(c(10, 20),"ft"), 
           self_treat = c(TRUE, FALSE)) # if self_treat is TRUE, a household cannot contaminate its own well

# Get probabilities
params_df$probs <- get_contamination_probabilities(params_df)
params_df[,c("density", "rs", "self_treat", "probs")]
```

